---
title: "LOL Figure 2"
author: "Eric Bridgeford"
date: "January 2, 2017"
output: html_document
---

# Figure 2 - LOL Paper

```{r, message=FALSE}
require(lol)
require(ggplot2)
require(latex2exp)
require(MASS)
require(gridExtra)

n <- 40
d <- 40
plot_sim <- function(X, Y, name) {
  data <- data.frame(x1=X[,1], x2=X[,2], y=Y)
  data$y <- factor(data$y)
  ggplot(data, aes(x=x1, y=x2, color=y)) +
    geom_point() +
    xlab(TeX("$x_1$")) +
    ylab(TeX("$x_2$")) +
    ggtitle(name) +
    labs(color="Class Label") +
    theme_bw()
}

# project into top 20 dimensions for X and Y and returns
# LDA projection plot
run_sim <- function(X, Y, func, ylab="", title="", r=1) {
  result <- do.call(func, list(X=X, Y=Y, r=r))
  
  liney <- lda(result$Xr, Y)
  result <- predict(liney, result$Xr)
  data <- data.frame(x1=result$x[,1], y=Y)
  data$y <- factor(data$y)
  proj_plot <- ggplot(data, aes(x=x1, fill=y, linetype=y)) +
    geom_density(alpha=0.6) +
    xlab(TeX("")) +
    ylab(TeX(ylab)) +
    theme_bw() +
    ggtitle(title) +
    scale_fill_discrete(name="Class Posterior") +
    guides(linetype=FALSE) +
    geom_jitter(data=data, aes(x=x1, y=0, group=y, color=y, shape=y), position = position_jitter(width=0, height=0.2))
  return(proj_plot)
}

g_legend<-function(a.gplot){
  tmp <- ggplot_gtable(ggplot_build(a.gplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)
}
```

In this notebook, we will consider 3 simulations with $K=2$ classes: Stacked Cigars, trunk, and rotated trunk. We will simulate $n=100$ examples in $d=1000$ dimensions, and will use LDA to estimate the $K-1=1$ dimensional projection optimizing the class discriminant boundary using `MASS::lda`.

We begin by generating the simulations:

```{r, message=FALSE}
sims <- list(lol.sims.cigar, lol.sims.rtrunk, lol.sims.rtrunk, lol.sims.mean_diff)
rotate <- c(FALSE, FALSE, TRUE, FALSE)
sim_names <-  c("Stacked Cigars", "Trunk", "Rotated Trunk", "Mean Difference")
funcs <- list(lol.project.pca, lol.project.cpca, lol.project.lrcca,
              lol.project.lol)
alg_names <- c("PCA", "PCA'", "LR-CCA", "LOL")
  
settings <- list(list(), list(), list(), list(md=4.4))

sim_plots <- list()
res_plots <- list()
counter1 <- 1; counter2 <- 1
for (i in 1:length(settings)) {
  data <- do.call(sims[[i]], c(list(n, d), settings[[i]]))
  sim_plots[[counter1]] <- plot_sim(data$X, data$Y, sim_names[i])
  for (j in 1:length(funcs)) {
    res_plots[[counter2]] <- run_sim(data$X, data$Y, funcs[[j]], ylab=alg_names[j],
                                     title=sim_names[i])
    counter2 <- counter2 + 1
  }
  counter1 <- counter1 + 1
}
```

Below, we plot the top column showing the simulation setup colored by class, showing the first 2 dimensions. The next 4 rows show the 2-class conditional posteriors resulting from projecting the data using the manifold technique indicated in the left-most column:

```{r, message=FALSE, fig.show='hide'}
top_legend <- g_legend(sim_plots[[1]])  # legend for the top most row
bottom_legend <- g_legend(res_plots[[length(sim_names) + 1]])  # legend for the rest of the examples
sim_plots <- sapply(sim_plots, function(simp) {
  simp + xlab("") + ylab("") + theme(legend.position=NaN)
  }, simplify=FALSE)

res_plots <- sapply(1:length(res_plots), function(j) {
    resp <- res_plots[[j]] + xlab("") + ggtitle("") +theme(legend.position=NaN)
    # remove the ylabel of only the non-left most columns
    if (j > length(sim_names)) {
      resp <- resp + ylab("")
    }
    return(resp)
  }, simplify=FALSE)
sim_grid <- grid.arrange(grobs=sim_plots, ncol=length(sim_names))
res_grid <- grid.arrange(grobs=res_plots, ncol=length(sim_names), as.table=FALSE)

res_bottom <- grid.arrange(res_grid, bottom_legend, nrow=1, widths=c(.88, .12))
res_top <- grid.arrange(sim_grid, top_legend, nrow=1, widths=c(.88, .12))
```

Plotting:

```{r, fig.height=10, fig.width=12}
ncol <- length(alg_names) + 1
fig2 <- grid.arrange(res_top, res_bottom, nrow=2,
                     heights=c(1/ncol, (ncol - 1)/ncol))
```
