---
title: "figure_5.Rmd"
author: "Eric Bridgeford"
date: "February 8, 2018"
output: html_document
---

```{r}
require(lolR)
require(ggplot2)
require(latex2exp)
require(MASS)
require(gridExtra)
require(data.table)
require(reshape2)
require(R.matlab)
require(grid)
require(plyr)
require(scales)

# compute the cutoff for the particular trial to get an approximate elbow
# by computing the smallest r with an associated lhat within 5%
# of the global minimum lhat
compute_cutoff <- function(rs, lhats, t=0.05) {
  sr.ix <- sort(rs, decreasing=FALSE, index.return=TRUE)$ix
  # compute minimum value
  min.lhat <- min(lhats)
  # compute minimum value + 5%
  lhat.thresh <- (1 + t)*min.lhat
  # find which indices are all below this
  lhat.below <- which(lhats <= lhat.thresh)
  rs.below <- rs[lhat.below]; lhats.below <- lhats[lhat.below]
  tmin.ix <- min(rs.below, index.return=TRUE)
  return(list(r=rs.below[tmin.ix], lhat=lhats.below[tmin.ix]))
}

g_legend<-function(a.gplot){
  tmp <- ggplot_gtable(ggplot_build(a.gplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)
}
```

```{r}
toep <- readMat('./data/fig3/toeplitz.mat')
tr2 <- readMat('./data/fig3/rtrunk.mat')
tr3 <- readMat('./data/fig3/3trunk.mat')
ft <- readMat('./data/fig3/fat_tails.mat')
qd <- readMat('./data/fig3/r2toeplitz.mat')

maxr <- c(90, 30, 30, 30, 30)
minr <- 0
mats <- list(toep, tr2, tr3, ft, qd)
sim.name <- c("Toeplitz", "Trunk-2", "Trunk-3", "Fat-Tails (D=1000)", "QDA")

interest <- list(c("ROAD"), c("ROAD"), c("LASSO"), c("ROAD"), c("ROAD"))
key <- c("ROAD", "lasso")
names(key) <- c("ROAD", "LASSO")


resultsm <- data.frame(sim=c(), iter=c(), alg=c(), r=c(), lhat=c())

for (k in 1:length(mats)) { 
  dat <- mats[[k]]
  desired_r <- 1:maxr[k]
  for (i in 1:length(dat$ks)) {  # i encodes simulation iteration
    for (j in length(interest[[k]])) {
      algname <- key[interest[[k]][j]]
      algid <- which(dimnames(dat$ks[[i]][[1]])[[1]] == algname)
      rs <- dat$ks[[i]][[1]][algid,,1][[algname]]
      algid <- which(dimnames(dat$Lhat)[[1]] == algname)
      lhats <- dat$Lhat[algid,,][[i]]
      lhat_adjust <- spline(rs, lhats, xout=desired_r, method='fmm', ties=mean)
      resultsm <- rbind(resultsm, data.frame(sim=sim.name[k], iter=i, alg=interest[[k]][j],
                                             r=lhat_adjust$x, lhat=lhat_adjust$y))
    }
  }
}
```

```{r}
maxr <- c(30, 90, 30, 30, 30)
ds <- c(100, 100, 100, 1000, 100)
# additional arguments for each simulation scenario
opt_args <- list(list(), list(), list(K=3), list(rotate=TRUE), list())
dat.names = c("Trunk-2", "Toeplitz", "Trunk-3", "Fat-Tails (D=1000)", "QDA")
titles <- c("Trunk-2", "Toeplitz", "Trunk-3", "Fat-Tails", "QDA")
dat.abbrs <- c("T", "Z", "3", "F", "Q")
names(dat.abbrs) <- dat.names

# read the results in
results <- readRDS('./data/fig3/lol_fig3_lda.rds')
results <- rbind(results$overall, resultsm)
#results <- results$overall
nan.mean <- function(x) mean(x, na.rm=TRUE)
results.means <- aggregate(lhat ~ sim + alg + r + lhat, data = results, FUN = nan.mean)
results.means$alg <- revalue(results.means$alg, c("cPCA"="LDA"))

acols <- c("#008000", "#4daf4a", "#e41a1c", "#ff7f00", "#377eb8", "#f781bf", "#00ffff")
algs <- c("LOL", "QOQ", "CCA", "ROAD", "LASSO", "PCA", "LDA")
names(acols) <- algs
```

```{r}
plot.results <- data.frame(r=c(), lhat=c(), sim=c(), alg=c())
for (i in 1:length(dat.names)) {
  for (j in 1:length(algs)) {
    alg <- algs[j]
    ss <- results.means[results.means$sim == dat.names[i] & results.means$alg == algs[j],]
    rs <- ss$r; lhats <- ss$lhat
    min.result <- compute_cutoff(rs, lhats)
    r.min <- min.result$r; lhat.min <- min.result$lhat
    if (alg == 'LOL') {
      norm.r <- r.min
      norm.lhat <- lhat.min
    }
    plot.results <- rbind(plot.results, data.frame(r=r.min/norm.r, lhat=lhat.min/norm.lhat,
                                                   sim=titles[i], alg=alg))
  }
}
```

```{r}
box <- data.frame(x=c(.1, 1, 1, .1), y=c(.1, .1, 1, 1))
panela <- ggplot(plot.results, aes(x=r, y=lhat)) +
  geom_polygon(data=box, aes(x=x, y=y), fill='gray', color='gray') +
  geom_point(aes(x=r, y=lhat, shape=sim, color=alg), size=3) +
  scale_color_manual(values=acols) +
  ylab("Normalized Misclassification Rate") +
  xlab("Normalized Embedding Dimension") +
  labs(shape="Simulation", color="Algorithm") +
  ggtitle("Simulated Data Performance") +
  scale_y_continuous(trans=log10_trans(), limits=c(.1, 10)) +
  scale_x_continuous(trans=log10_trans(), limits=c(.1, 10)) +
  theme_bw() +
  theme(legend.position=NaN)
panela_leg <- ggplot(plot.results, aes(x=r, y=lhat)) +
  geom_polygon(data=box, aes(x=x, y=y), fill='gray', color='gray') +
  geom_point(aes(x=r, y=lhat, shape=sim), size=3) +
  scale_color_manual(values=acols) +
  ylab("Normalized Misclassification Rate") +
  xlab("Normalized Embedding Dimension") +
  labs(shape="Simulation", color="Algorithm") +
  ggtitle("Simulated Data Performance") +
  scale_y_continuous(trans=log10_trans(), limits=c(.1, 10)) +
  scale_x_continuous(trans=log10_trans(), limits=c(.1, 10)) +
  theme_bw()
leg_panela <- g_legend(panela_leg)
```


---
title: "LOL Figure 3 with LDA Classifier"
author: "Eric Bridgeford"
date: "January 5, 2017"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Figure 3 - LOL Paper

```{r, message=FALSE}

# compute the cutoff for the particular trial to get an approximate elbow
# by computing the smallest r with an associated lhat within 5%
# of the global minimum lhat
compute_cutoff <- function(rs, lhats, t=0.05) {
  sr.ix <- sort(rs, decreasing=FALSE, index.return=TRUE)$ix
  # compute minimum value
  min.lhat <- min(lhats)
  # compute minimum value + 5%
  lhat.thresh <- (1 + t)*min.lhat
  # find which indices are all below this
  lhat.below <- which(lhats <= lhat.thresh)
  rs.below <- rs[lhat.below]; lhats.below <- lhats[lhat.below]
  tmin.ix <- min(rs.below, index.return=TRUE)
  return(list(r=rs.below[tmin.ix], lhat=lhats.below[tmin.ix]))
}


plot_sim_lhats <- function(data, cols, pt.dat, title="", by=10, from=10, ylab=TeX("$\\hat{L}$"),
                           xlab="Embedded Dimensions", fsize=12) {
  lims <- c(floor(10*min(data$lhat))/10, ceiling(10*max(data$lhat))/10)
  breaks= seq(from=lims[1], to=lims[2], by=0.1)
  xlims <- c(min(data$r), max(data$r))
  xbreaks <- seq(from=from, to=xlims[2], by=by)
  plot_sims <- ggplot(data, aes(x=r, y=lhat, group=alg, color=alg)) +
    geom_line(data=base::subset(data, alg == "CCA"), aes(x=r, y=lhat, group=alg, color=alg), size=.75) +
    geom_point(data=base::subset(pt.dat, alg == "CCA"), aes(x=r, y=lhat, group=alg, color=alg), size=2) +
    geom_line(data=base::subset(data, alg != "CCA" & alg != "QOQ"), aes(x=r, y=lhat, group=alg, color=alg), size=.75) +
    geom_point(data=base::subset(pt.dat, alg != "CCA"), aes(x=r, y=lhat, group=alg, color=alg), size=2) +
    geom_line(data=base::subset(data, alg == "QOQ"), aes(x=r, y=lhat, group=alg, color=alg), linetype="dashed", size=.75) +
    xlab(xlab) +
    ylab(ylab) +
    ggtitle(title) +
    theme_bw() +
    scale_y_continuous(limits=lims, breaks=breaks) +
    scale_x_continuous(limits=xlims, breaks=xbreaks) +
    scale_color_manual(values=cols, limits=names(cols),
                       guide=guide_legend(ncol=1)) +#, title.position="top", title.hjust = .5)) +
    labs(color="Algorithm", linetype="Test")
  return(plot_sims)
}

g_legend<-function(a.gplot){
  tmp <- ggplot_gtable(ggplot_build(a.gplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)
}
```

The below code will produce the required data, which runs LOL, cPCA, PCA, and LR-CCA at the desired simulation settings. Note that this function will multi-thread natively, and took approximately 7 hours to run on a 96 core machine with $\frac{7}{8}$ of the cores active.

```{r, eval=FALSE}
source('./figure_4_driver.R')
```

Borrowing results from an earlier matlab implementation:

```{r}
col <- readMat('./data/colon/Colon.mat')
mn <- readMat('./data/mnist/MNIST.mat')

maxr <- c(20, 30)
minr <- 1
mats <- list(col, mn)
exp_name <- c("Colon", "MNIST")

interest <- list(c("ROAD", "LASSO"), c("LASSO"))
key <- c("ROAD", "lasso")
names(key) <- c("ROAD", "LASSO")

resultsm <- data.frame(data=c(), iter=c(), alg=c(), r=c(), lhat=c())

for (k in 1:length(mats)) { 
  dat <- mats[[k]]
  desired_r <- 1:maxr[k]
  for (i in 1:length(dat$ks)) {  # i encodes simulation iteration
    for (j in length(interest[[k]])) {
      algname <- key[interest[[k]][j]]
      algid <- which(dimnames(dat$ks[[i]][[1]])[[1]] == algname)
      rs <- dat$ks[[i]][[1]][algid,,1][[algname]]
      algid <- which(dimnames(dat$Lhat)[[1]] == algname)
      lhats <- dat$Lhat[algid,,][[i]]
      lhat_adjust <- spline(rs, lhats, xout=desired_r, method='fmm', ties=mean)
      resultsm <- rbind(resultsm, data.frame(data=exp_name[k], iter=i, alg=interest[[k]][j],
                                             r=lhat_adjust$x, lhat=lhat_adjust$y))
    }
  }
}

n <- c(62, 1000)
names(n) <- c("Colon", "MNIST")
resultsm <- data.table(resultsm)
resultsm.means <- aggregate(lhat ~ alg + r + data, data=resultsm, function(x) c(lhat=mean(x), sd=sd(x)))
resultsm.means <- cbind(resultsm.means, se=resultsm.means$lhat[,"sd"]/sqrt(n[match(resultsm.means$data, names(n))]))
resultsm.means <- data.frame(alg=resultsm.means$alg, r=resultsm.means$r, data=resultsm.means$data, lhat=resultsm.means$lhat[,"lhat"], se=resultsm.means$se)
```

```{r}
exp_names = c("Colon", "MNIST")
titles <- c("Colon, n=62, d=2000", "MNIST, n=1000, d=784")
exp_min <- c(10, 10)
by <- c(10, 10)
```

```{r}
# read the results in
results <- readRDS('./data/fig4/lol_fig4_lda.rds')
results.means <- rbind(results, resultsm.means)
results.means$alg <- revalue(results.means$alg, c("cPCA"="LDA"))
#results <- results$overall

acols <- c("#008000", "#4daf4a", "#bcbddc", "#9e9ac8", "#807dba", "#6a51a3", "#4a1486")
names(acols) <- c("LOL", "QOQ", "CCA", "ROAD", "LASSO", "PCA", "LDA")
```

```{r}
plot.results <- data.frame(r=c(), lhat=c(), exp=c(), alg=c())
for (i in 1:length(exp_names)) {
  for (j in 1:length(algs)) {
    alg <- algs[j]
    ss <- results.means[results.means$data == exp_names[i] & results.means$alg == algs[j],]
    rs <- ss$r; lhats <- ss$lhat
    min.result <- compute_cutoff(rs, lhats)
    r.min <- min.result$r; lhat.min <- min.result$lhat
    if (alg == 'LOL') {
      norm.r <- r.min
      norm.lhat <- lhat.min
    }
    plot.results <- rbind(plot.results, data.frame(r=r.min/norm.r, lhat=lhat.min/norm.lhat,
                                                   exp=exp_names[i], alg=alg))
  }
}
```

```{r}
plot.results$exp <- factor(plot.results$exp)
box <- data.frame(x=c(.1, 1, 1, .1), y=c(.1, .1, 1, 1))
panelb <- ggplot(plot.results, aes(x=r, y=lhat)) +
  geom_polygon(data=box, aes(x=x, y=y), fill='gray', color='gray') +
  geom_point(aes(x=r, y=lhat, shape=exp, color=alg), size=3) +
  scale_color_manual(values=acols) +
  ylab("Normalized Misclassification Rate") +
  xlab("Normalized Embedding Dimension") +
  labs(color="Algorithm", shape="Experiment") +
  ggtitle("Real Data Performance") +
  scale_y_continuous(trans=log10_trans(), limits=c(.1, 10)) +
  scale_x_continuous(trans=log10_trans(), limits=c(.1, 10)) +
  theme_bw() +
  theme(legend.position=NaN)
panelb_leg <- ggplot(plot.results, aes(x=r, y=lhat)) +
  geom_polygon(data=box, aes(x=x, y=y), fill='gray', color='gray') +
  geom_point(aes(x=r, y=lhat, shape=exp), size=3) +
  ylab("Normalized Misclassification Rate") +
  xlab("Normalized Embedding Dimension") +
  labs(color="Algorithm", shape="Experiment") +
  ggtitle("Comparison of Embedding Techniques to LOL") +
  scale_y_continuous(trans=log10_trans(), limits=c(.1, 10)) +
  scale_x_continuous(trans=log10_trans(), limits=c(.1, 10)) +
  theme_bw()
leg_panelb <- g_legend(panelb_leg)
```

```{r, fig.width=10, fig.height=3}
overall_leg <- ggplot(plot.results, aes(x=r, y=lhat, color=alg)) +
  geom_polygon(data=box, aes(x=x, y=y), fill='gray', color='gray') +
  geom_point(size=3) +
  scale_color_manual(values=acols) +
  ylab("Normalized Misclassification Rate") +
  xlab("Normalized Embedding Dimension") +
  labs(color="Algorithm", shape="Experiment") +
  ggtitle("Comparison of Embedding Techniques to LOL") +
  scale_y_continuous(trans=log10_trans(), limits=c(.1, 10)) +
  scale_x_continuous(trans=log10_trans(), limits=c(.1, 10)) +
  theme_bw()
leg_overall <- g_legend(overall_leg)
grid.newpage()
grid.arrange(arrangeGrob(panela, leg_panela, widths=c(.8, .25)),
             arrangeGrob(panelb + xlab("") + ylab(""), leg_panelb, widths=c(.8, .25)),
             leg_overall, widths=c(.45, .45, .1), ncol=3)
```
