---
title: "LOL Cross-Validation"
author: "Eric Bridgeford"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{xval}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Cross Validation to Determine Error of Embedding Alg composed with Classifier

```{r}
require(lolR)
require(ggplot2)
n = 400
d = 30
r = 3
```

Here, we look at how to see the generalizability of a given model in the form of the cross-validated error. We simulate data with `n=200` and `d=30`:

```{r, fig.width=5}
testdat <- lol.sims.rtrunk(n, d)
X <- testdat$X
Y <- testdat$Y

data <- data.frame(x1=X[,1], x2=X[,2], y=Y)
data$y <- factor(data$y)
ggplot(data, aes(x=x1, y=x2, color=y)) +
  geom_point() +
  xlab("x1") +
  ylab("x2") +
  ggtitle("Simulated Data")
```

We arbitrarily select LOL as our algorithm, and look at the leave-one-out (loo) cross-validated error with the LDA classifier. We project the resulting model to `3` dimensions and visualize the first `2`:

```{r, fig.width=5}
result <- lol.xval.eval(X, Y, r, alg = lol.project.lol, alg.return="A",
                        classifier=MASS::lda, classifier.return="class", k='loo')

data <- data.frame(x1=result$model$Xr[,1], x2=result$model$Xr[,2], y=Y)
data$y <- factor(data$y)
ggplot(data, aes(x=x1, y=x2, color=y)) +
  geom_point() +
  xlab("x1") +
  ylab("x2") +
  ggtitle(sprintf("Projected Data using LOL, L=%.2f", result$lhat))
```

# Cross Validation to Determine Optimal Embedding Dimensions of Embedding Alg composed with Classifier

```{r}
result <- lol.xval.optimal_dimselect(X, Y, rs=c(5, 10, 15), alg = lol.project.lol, alg.return="A",
                        classifier=MASS::lda, classifier.return="class", k='loo')

data <- data.frame(x1=result$model$Xr[,1], x2=result$model$Xr[,2], y=Y)
data$y <- factor(data$y)
ggplot(data, aes(x=x1, y=x2, color=y)) +
  geom_point() +
  xlab("x1") +
  ylab("x2") +
  ggtitle(sprintf("Projected Data using LOL, L=%.2f", result$optimal.lhat))

ggplot(result$foldmeans.data, aes(x=r, y=lhat)) +
  geom_line() +
  xlab("Embedding Dimensions, r") +
  ylab("Misclassification Rate, L") +
  ggtitle("Impact on Misclassification Rate of Embedding Dimension")

print(sprintf("optimal dimension: %d", result$optimal.r))
print(sprintf("Misclassification rate at rhat = %d: %.2f", result$optimal.r, result$optimal.lhat))
```
