---
title: "Check QO Code"
output:
  html_document:
    df_print: paged
---

```{r, message=FALSE}
require(lolR)
require(ggplot2)
require(MASS)
require(abind)
n=1000
d=5
r=3
rot=-pi/4  # rotation in radians
R = rbind(c(cos(rot), -sin(rot), 0, 0, 0), c(sin(rot), cos(rot), 0, 0, 0),
          c(0,0,1,0,0), c(0,0,0,1,0), c(0,0,0,0,1))  # rotation matrix for rotation by 45 degrees in first 2 dimensions

lol.project.qoq <- function(X, Y, r, xfm=FALSE, xfm.opts=list(), ...) {
  # class data
  info <- lolR:::lol.utils.info(X, Y)
  priors <- info$priors; centroids <- info$centroids
  K <- info$K; ylabs <- info$ylabs
  n <- info$n; d <- info$d
  if (r > d) {
    stop(sprintf("The number of embedding dimensions, r=%d, must be lower than the number of native dimensions, d=%d", r, d))
  }
  deltas <- lolR:::lol.utils.deltas(centroids, priors)[, 2:K]
  centroids <- t(centroids)

  nv <- r - (K) + 1
  Aclass <- array(0, dim=c(d, 0))  # the class-wise egvecs
  vclass <- c()  # the class-wise egvals
  vclass.res <- list(d=NULL)
  if (nv >= 0) {
    for (ylab in ylabs) {
      Xclass = X[Y == ylab,]
      obj <- lol.project.pca(Xclass, r=nv, xfm=xfm, xfm.opts=xfm.opts)
      Aclass <- cbind(Aclass, obj$A)
      vclass <- c(vclass, obj$d[1:nv])
    }
    # take the nv from the A computed for each class using the
    # nv with the top eigenvalues from Aclass
    A <- cbind(deltas, Aclass[, sort(vclass, index.return=TRUE, decreasing=TRUE)$ix[1:nv]])
    vclass.res$d <- sort(vclass)[1:nv]
  } else {
    A <- deltas[, 1:r, drop=FALSE]
  }
  colnames(A) <- NULL

  return(list(A=A, d=vclass.res$d, centroids=centroids, priors=priors, ylabs=ylabs,
              Xr=lol.embed(X, A), cr=lol.embed(centroids, A)))
}

lol.project.lol <- function(X, Y, r, xfm=FALSE, xfm.opts=list(), ...) {
  # class data
  info <- lolR:::lol.utils.info(X, Y)
  priors <- info$priors; centroids <- info$centroids
  K <- info$K; ylabs <- info$ylabs
  n <- info$n; d <- info$d
  if (r > d) {
    stop(sprintf("The number of embedding dimensions, r=%d, must be lower than the number of native dimensions, d=%d", r, d))
  }
  deltas <- lolR:::lol.utils.deltas(centroids, priors)[, 2:K]
  centroids <- t(centroids)

  nv <- r - (K) + 1
  lrlda <- list(d=NULL)
  if (nv > 0) {
    lrlda <- lol.project.lrlda(X, Y, r=nv, xfm=xfm, xfm.opts=xfm.opts)
    A <- cbind(deltas, lrlda$A)
  } else {
    A <- deltas[, 1:r, drop=FALSE]
  }
  colnames(A) <- NULL

  # orthogonalize and normalize
  A <- qr.Q(qr(A))
  return(list(A=A, d=lrlda$d, centroids=centroids, priors=priors, ylabs=ylabs,
              Xr=lol.embed(X, A), cr=lol.embed(centroids, A)))
}
```

# Generate Data

```{r}
mu0 = array(0, dim=c(d))
Sigma0 = diag(d)
Sigma0[1,1] = 5
mu1 = t(R) %*% mu0
Sigma1 = t(R) %*% Sigma0 %*% R

mus = abind(mu0, mu1, along=2)
Sigmas = abind(Sigma0, Sigma1, along=3)
data = lolR:::lol.sims.sim_gmm(mus, Sigmas, n=n, priors = c(0.5, 0.5))
X = data$X; Y = data$Y
```

# Visualize Data

Below, we visualize the data rotated 45 degrees in the first and second dimensions:

```{r}
X.dat = data.frame(x1=X[,1], x2=X[,2], class=Y)
X.dat$class <- factor(X.dat$class)
ggplot(X.dat, aes(x=x1, y=x2, color=class, group=class, fill=class)) +
  geom_point() + 
  stat_ellipse(type = "norm", level=.6825)
```

# Projection vectors using LO

We look at the second 2 vectors the data projects onto with LOL (the top 2 non-mean vectors). These should be "between" the two ellipses:

```{r}
lol.result <- lol.project.lol(X, Y, r=4)
proj1.result <- data.frame(x1=rbind(lol.result$A[1,2], 0), x2=rbind(lol.result$A[2,2], 0), vector=1)
proj2.result <- data.frame(x1=rbind(0, lol.result$A[1,3]), x2=rbind(0, lol.result$A[2,3]), vector=2)
proj.dat = rbind(proj1.result, proj2.result)
proj.dat$vector <- factor(proj.dat$vector)
ggplot(X.dat, aes(x=x1, y=x2, color=class, group=class)) +
  geom_point() + 
  stat_ellipse(type = "norm", level=.6825) +
  geom_line(data=proj.dat, aes(x=x1, y=x2, color=vector, group=vector), color='black')
```

As we can see, the directions are orthogonal to the point clouds due to the class-conditional covariance structure.

# Projection vectors using QO

```{r}
lol.result <- lol.project.qoq(X, Y, r=4)
proj1.result <- data.frame(x1=rbind(lol.result$A[1,2], 0), x2=rbind(lol.result$A[2,2], 0), vector=1)
proj2.result <- data.frame(x1=rbind(0, lol.result$A[1,3]), x2=rbind(0, lol.result$A[2,3]), vector=2)
proj.dat = rbind(proj1.result, proj2.result)
proj.dat$vector <- factor(proj.dat$vector)
ggplot(X.dat, aes(x=x1, y=x2, color=class, group=class)) +
  geom_point() + 
  stat_ellipse(type = "norm", level=.6825) +
  geom_line(data=proj.dat, aes(x=x1, y=x2, color=vector, group=vector), color='black')
```

as we can see, the directions are exactly in-line with the point clouds.
