---
title: "Check QO Code"
output:
  html_document:
    df_print: paged
---

```{r, message=FALSE}
require(lolR)
require(ggplot2)
require(MASS)
require(abind)
n=1000
d=5
r=3
rot=-pi/4  # rotation in radians
R = rbind(c(cos(rot), -sin(rot), 0, 0, 0), c(sin(rot), cos(rot), 0, 0, 0),
          c(0,0,1,0,0), c(0,0,0,1,0), c(0,0,0,0,1))  # rotation matrix for rotation by 45 degrees in first 2 dimensions
```

# Generate Data

```{r}
mu0 = array(0, dim=c(d))
Sigma0 = diag(d)
Sigma0[1,1] = 5
mu1 = t(R) %*% mu0
Sigma1 = t(R) %*% Sigma0 %*% R

mus = abind(mu0, mu1, along=2)
Sigmas = abind(Sigma0, Sigma1, along=3)
data = lolR:::lol.sims.sim_gmm(mus, Sigmas, n=n, priors = c(0.5, 0.5))
X = data$X; Y = data$Y
```

# Visualize Data

Below, we visualize the data rotated 45 degrees in the first and second dimensions:

```{r}
X.dat = data.frame(x1=X[,1], x2=X[,2], class=Y)
X.dat$class <- factor(X.dat$class)
ggplot(X.dat, aes(x=x1, y=x2, color=class, group=class, fill=class)) +
  geom_point() + 
  stat_ellipse(type = "norm", level=.6825)
```

# Projection vectors using LO

We look at the second 2 vectors the data projects onto with LOL (the top 2 non-mean vectors). These should be "between" the two ellipses:

```{r}
lol.result <- lol.project.lol(X, Y, r=4, orthogonalize=FALSE)
proj1.result <- data.frame(x1=rbind(lol.result$A[1,2], 0), x2=rbind(lol.result$A[2,2], 0), vector=1)
proj2.result <- data.frame(x1=rbind(0, lol.result$A[1,3]), x2=rbind(0, lol.result$A[2,3]), vector=2)
proj.dat = rbind(proj1.result, proj2.result)
proj.dat$vector <- factor(proj.dat$vector)
ggplot(X.dat, aes(x=x1, y=x2, color=class, group=class)) +
  geom_point() + 
  stat_ellipse(type = "norm", level=.6825) +
  geom_line(data=proj.dat, aes(x=x1, y=x2, color=vector, group=vector), color='black')
```

As we can see, the directions are orthogonal to the point clouds due to the class-conditional covariance structure.

# Projection vectors using QO

```{r}
lol.result <- lol.project.lol(X, Y, r=4, second.moment='quadratic', orthogonalize=FALSE)
proj1.result <- data.frame(x1=rbind(lol.result$A[1,2], 0), x2=rbind(lol.result$A[2,2], 0), vector=1)
proj2.result <- data.frame(x1=rbind(0, lol.result$A[1,3]), x2=rbind(0, lol.result$A[2,3]), vector=2)
proj.dat = rbind(proj1.result, proj2.result)
proj.dat$vector <- factor(proj.dat$vector)
ggplot(X.dat, aes(x=x1, y=x2, color=class, group=class)) +
  geom_point() + 
  stat_ellipse(type = "norm", level=.6825) +
  geom_line(data=proj.dat, aes(x=x1, y=x2, color=vector, group=vector), color='black')
```

as we can see, the directions are exactly in-line with the point clouds.
